Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-10-03 arxiv/proof_examples.tex
revise

		 g(A) \define & \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)+ \left(A^2 X^{(1)}\Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} + A X^{(1)}\Sigma^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
		 + & \left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)+ \left(X^{(2)}\Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)\epsilon^{(2)} + A X^{(2)}\Sigma^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2, 

@@

	$$\Sigma\equiv \Sigma(A):= A^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$

@@

&= \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon_1} +\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2\epsilon^{(2)}}   \\
&+  A^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)} \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\Sigma^{-1} (X^{(2)})^\top \epsilon^{(2)}}  +A^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)} \Sigma^{-1}  (X^{(2)})^\top X^{(2)}\Sigma^{-1} (X^{(1)})^\top \epsilon^{(1)}  }\\
		&= \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
		&+ \sigma^2\tr\left[ \left(A^2 X^{(1)}\Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\right] +\sigma^2\tr\left[ \left(X^{(2)}\Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2 \right]\\
&+ 2 A^2 \sigma^2 \tr\left[ \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\Sigma^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\
		& = \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)\right\|^2 +A^2\left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-A\beta_2)\right\|^2 +\sigma^2(n_1+n_2-p).

@@

&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right] \nonumber\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p) \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}

@@

where in the first step we used that the entries of $\beta_1-A\beta_2\in \R^p$ are i.i.d. random variables with mean zero and variance $p^{-1}[(A-1)^2\kappa^2 + (A^2+1)d^2/2]$, and in the second step we used  %Using $(X^{(2)})^\top X^{(2)}=\hat\Sigma- A^2 (X^{(1)})^\top X^{(1)}$, we can further simplify that

@@

 &\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] \\
 &=\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} \right] \\
 & = \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right] .

@@

 In the first step of the above derivation, we used that 
 $$(X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} = (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)}.$$

Using the concentration estimates in Lemma \ref{largedeviation}, we can control the difference between $g(A)$ and its partial expectation. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(A) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
$$\left| g(A) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2 }\right| \le p^{-1/2+e_1}\cdot p\left(\sigma^2 + \kappa^2+d^2 \right) $$
for any small constant $e_1>0$.
 \end{claim}
 \begin{proof}
 \todo{add proof}
 \end{proof}

>>>>

		 g(A) \define & \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)+ \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} + A X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
		 + & \left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)+ \left(X^{(2)}\Sigma(A)^{-1} (X^{(2)})^\top-\id_{n_2}\right)\epsilon^{(2)} + A X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2, 

@@

	$$ \Sigma(A):= A^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$

@@

&= \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon^{(1)}} +\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\Sigma(A)^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2\epsilon^{(2)}}   \\
&+  A^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)} \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)}}  +A^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)} \Sigma(A)^{-1}  (X^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top \epsilon^{(1)}  }\\
		&= \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
		&+ \sigma^2\tr\left[ \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\right] +\sigma^2\tr\left[ \left(X^{(2)}\Sigma(A)^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2 \right]\\
&+ 2 A^2 \sigma^2 \tr\left[ \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\
		& = \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)\right\|^2 +A^2\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-A\beta_2)\right\|^2 +\sigma^2(n_1+n_2-p).

@@

&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)}\right] \nonumber\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p) \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}

@@

where in the first step we used that the entries of $\beta_1-A\beta_2\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(A-1)^2\kappa^2 + (A^2+1)d^2/2]$, and in the second step we used  %Using $(X^{(2)})^\top X^{(2)}=\hat\Sigma- A^2 (X^{(1)})^\top X^{(1)}$, we can further simplify that

@@

 & (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} + A^2   (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)}  \\
 &=  (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} + A^2  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}  \\
 & =   (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} .

@@

Here in the first step of the above derivation, we used that 
 $$(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} = (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)}.$$

Using the concentration estimates in Lemma \ref{largedeviation}, we can control the difference between $g(A)$ and its partial expectation. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(A) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
$$\left| g(A) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2 }\right| \le p^{1/2+e_1} \left(\sigma^2 + \kappa^2+d^2 \right) $$
for any small constant $e_1>0$.
 \end{claim}
 \begin{proof}
By Fact \ref{lem_minv} (iii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, 
 \ee
 and
 \be\label{op_Sig1}
\| \Sigma(A)^{-1}\|\le \frac1{A^2((\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}})+((\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}) }\lesssim \frac{1}{(A^2+1)p},
 \ee
 where we also used that $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$ for a small constant $\tau>0$.

 Now we expand the first term on the right-hand side of \eqref{eq_mtl_A12}: 
 \begin{align*}
 & \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)+ \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} + A X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
 &=h_1(A) + h_2(A) +h_3(A) + 2h_4(A) + 2h_5(A) + 2h_6(A) ,
 \end{align*}
 where
 \begin{align*}
&h_1(A):=  (\beta_1-A\beta_2)^\top (X^{(2)})^\top X^{(2)}  \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)  ,\\
& h_2(A) := {(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon_1}  ,\\
& h_3(A):=  A^2 { (\epsilon^{(2)} )^\top X^{(2)} \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)}} , \\
& h_4(A):=  (\epsilon^{(1)})^\top\left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right) X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1), \\
&h_5(A):=A (\epsilon^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1),\\
&h_6(A):=A (\epsilon^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top\left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)}.
\end{align*}
Next we estimate each term using Lemma \ref{largedeviation} in Appendix \ref{app_tool}.

For $h_1(A)$, using Lemma \ref{largedeviation} in Appendix \ref{app_tool} and the fact that the entries of $\beta_1-A\beta_2\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(A-1)^2\kappa^2 + (A^2+1)d^2/2]$, we obtain the following estimate with high probability:
\begin{align*}
&\left|h_1(A) - \exarg{\beta_1,\beta_2}{h_1(A) \mid X_1, X_2}\right| \\
&\le p^{\e_1}\cdot p^{-1}[(A-1)^2\kappa^2 + (A^2+1)d^2/2]\cdot \left\| (X^{(2)})^\top X^{(2)}  \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} \right\|_F \\
&\le p^{e_1}\cdot p^{-1}[(A-1)^2\kappa^2 + (A^2+1)d^2/2]\cdot p^{1/2} \left\| (X^{(2)})^\top X^{(2)}  \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} \right\|  \\
&\lesssim p^{1/2+e_1}\cdot \frac{(A-1)^2\kappa^2 + (A^2+1)d^2/2}{(A^2+1)^2}\lesssim p^{1/2+e_1}(\kappa^2 + d^2),
\end{align*}
for any small constant $\e_1>0$. Here in the last step we used \eqref{op_X12} and \eqref{op_Sig1} to bound the operator norm:
\begin{align*}
&\left\| (X^{(2)})^\top X^{(2)}  \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} \right\| \\
&\le \left\| (X^{(2)})^\top X^{(2)}\right\|^2 \cdot \left\| (X^{(1)})^\top X^{(1)}\right\|\cdot \left\| \Sigma(A)^{-1} \right\|^2 \lesssim \frac{p}{(A^2+1)^2}.
\end{align*}

### 2020-10-03 arxiv/proof_same_cov.tex
revise

	For equation \eqref{eq_gA_p2}, using Lemma \ref{largedeviation} in Appendix \ref{sec_maintools} and the fact that all moments of $\varepsilon_i$ exist by Assumption \eqref{assmAhigh2}, we obtain that the following holds with high probability:
	\begin{align}
		|\inner{XB^{\star} (U_AU_A^{\top} - \id)}{\cE}| &\le p^\e \cdot \sigma \cdot \bignormFro{XB^{\star} (U_AU_A^{\top} - \id)} \nonumber \\
		&\le p^{\e} \cdot \sigma \cdot \norm{Z} \cdot \bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id)} \nonumber \\
		&\le p^{1/2 + \e} \cdot \sigma \cdot \bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id)}, \label{eq_gA_err2}
	\end{align}
	which is less than the right hand side of equation \eqref{eq_gA_err} by Cauchy-Schwartz inequality.
	In the second step, we use the fact that $X=Z\Sigma^{1/2}$.
	In the third step, we use Fact \ref{lem_minv} to bound the operator norm $\|Z\|$ by $\OO(\sqrt{p})$.

>>>>

	For equation \eqref{eq_gA_p2}, using Lemma \ref{largedeviation} in Appendix \ref{app_tool} and the fact that all moments of $\varepsilon_i$ exist by Assumption \eqref{assmAhigh2}, we obtain that the following holds with high probability:
	\begin{align}
		|\inner{XB^{\star} (U_AU_A^{\top} - \id)}{\cE}| &\le p^\e \cdot \sigma \cdot \bignormFro{XB^{\star} (U_AU_A^{\top} - \id)} \nonumber \\
		&\le p^{\e} \cdot \sigma \cdot \norm{Z} \cdot \bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id)} \nonumber \\
		&\le p^{1/2 + \e} \cdot \sigma \cdot \bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id)}, \label{eq_gA_err2}
	\end{align}
	which is less than the right hand side of equation \eqref{eq_gA_err} by Cauchy-Schwartz inequality.
	In the second step, we use the fact that $X=Z\Sigma^{1/2}$.
	In the third step, we use Fact \ref{lem_minv} to bound the operator norm $\|Z\|$ by $\OO(\sqrt{p})$.
